---
title: "Master_data_science_in_R"
output: html_document
author: Shamim Rashid
date: 13th november,2019
---

MASTER DATA SCIENCE IN R
```{r}
library(tidyverse)

```
EXERCISE 3.2.4

1. Run ggplot(data = mpg). What do you see?

-You see a blank plot since no x and y axis have been mapped.
```{r}
ggplot(data = mpg)
```

2.How many rows are in mpg?

-234 rows and 11 variables

3.What does the drv variable describe? Read the help for ?mpg to find out.

-drv- front wheel drive, rear wheel drive

4.Make a scatterplot of hwy vs cyl.

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=hwy ,y=cyl))
```

5.What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

The plot is not useful since we are mapping class vs drv yet they are both categorical data

```{R}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=class,y=drv))

```



EXERCISE 3.3.1

1.What’s gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

The points are not blue as the color attribute has to be set manually outside the aeshetics brackets.

Correction:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy),color = "blue")
```


2.Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

```{r}
mpg

```


continuous variables: disp, cyl, hwy, cty


categorica variables:class, fl, drv, trans, year



3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?



```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=hwy,y=cyl,color=hwy,shape=class,size=cyl));
```


For continous variables ,in the legend, values are shown while for categorical variables,names of the categories are indicated


4.What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=hwy,y=cyl,color= hwy, shape= class));

```



Only 6 rows are mapped on the plot as it is hard to discrimanate between hence a maximum of 6 discrete vaues are used



5.What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)


The stroke aesthetic only works with bordered shapes and is used to manipulate the width of the border.
```{r}

ggplot(mtcars, aes(wt, mpg)) +
  geom_point(shape = 21, colour = "black", fill = "white", size = 5, stroke = 5)


```



6.What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=hwy,y=displ,color=displ<5))
```




EXERCISE 3.5.1


1.What happens if you facet on a continuous variable?


Each value has a facet
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=hwy,y=cyl))+
  facet_grid(.~displ)
```



2.What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```


The empty cells in the plot means that there are no cars with the combination given hence the empty facets



3.What plots does the following code make? What does . do?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```


The dot attribute plots without the row attribute hence the y axis values are repeated.


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~drv)
```


The dot attribute plots without the column attribute hence the x axis values are repeated.



4.Take the first faceted plot in this section:What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?


(i)
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

With faceting,you are able to see individual classes clearly.


(ii)
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy,color=class))
```

With coloring ,you are able to see how classes are clustered all over.However, with larger datasets, individual classes are harder to see than clustered classes.



5.Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

```{r}
#nrow-specifies no. of rows required.
#ncol-specifies no.of columns required.
```

facet_grid() have no nrow and ncol arguments as the aesthetics attributes specify the number of rows and columns.


6.When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

if variables with more levels are placed on the column, y-axis would shrink making it harder to see points on the plot.